Expert-In-The-Loop Causal Discovery: Iterative Model Refinement Using Expert Knowledge

### 2024-08-07 paper/main.tex
updates

\begin{abstract}
	\emph{Causal Discovery} has received significant attention in the
	\emph{Directed Acyclic Graphs} (DAGs) literature, leading to the
	development of numerous automated algorithms for learning DAGs from
	data. Despite these advancements, the application of these algorithms
	in applied domains remain limited, with researchers often preferring to
	construct DAGs manually based on their expert knowledge. We hypothesize
	that the reluctance to use these algorithms arises due to several
	practical issues, such as algorithms often make obvious mistakes making
	it difficult to trust, they output the Markov Equivalence Class (MEC)
	instead of a DAG (which is often required for further analysis), and
	difficulties in choosing the appropriate algorithm and hyperparameters
	for a given dataset. Even if using these algorithms, researchers have
	to rely on their expert knowledge to make modifications to the learned
	model in terms of fixing any mistakes or orienting the edges before
	these models can be used for any further analysis. Given this reliance
	on manual model modification, in this paper, we present a method
	inspired by \emph{specification search} in \emph{Structural Equation
	Models} (SEMs) to assist researchers in iteratively modifying or
	constructing their models to improve their fit to data. Specifically,
	we utilize implied Conditional Independence (CI) based model testing
	combined with a novel measure of partial association for mixed data to
	rank violations of implied CIs of the model in data. Based on this
	ranking, users can then choose to make modifications to the model by
	either removing or adding edges while using their expertise to decide
	the orientation of the edge. The ranking allows users to prioritize the
	worst offending violations first. Through empirical analysis, when
	applied to constructing the model from scratch, we demonstrate that
	taking a greedy approach (i.e., fixing the worst violation first) gets
	us comparable performance to automated algorithms if the user is able
	to correctly provide the edge direction in $ 1 $ out of $ 3 $ prompts.
	We further show that in the absence of an expert, a Large Language
	Model (LLM) can also be used for getting the edge orientations.
\end{abstract}

>>>>

\begin{abstract}
Causal discovery has received significant attention in the Directed Acyclic
Graphs (DAGs) literature, leading to the development of numerous automated
algorithms for learning DAGs from data. Despite these advancements, the
application of these algorithms in applied domains remains limited, with
researchers often preferring to manually construct DAGs based on their expert
knowledge. Even when using these algorithms, researchers must rely on their
expertise to correct any mistakes or orient edges before the models can be used
for further analysis. Given this reliance on manual model construction and
modification, we propose a method inspired by specification search in
Structural Equation Models (SEMs) to assist researchers in iteratively
modifying DAGs. Our method employs implied Conditional Independence (CI) based
model testing, combined with a novel measure of partial association for mixed
data, to rank violations of implied CIs of the model in the data. This ranking
helps the user in choosing pairs of variables between which to add or remove
edges. Based of their expertise, the user can decide to remove or add edge
while deciding the correct orientation between the pair of variables.
Empirically, we demonstrate that a greedy approach (i.e., fixing the worst
violation first) achieves performance comparable to automated algorithms,
provided the user can correctly determine the edge direction in $ 1 $ out of $
3 $ edge additions. Additionally, we show that in the absence of an expert, a
Large Language Model (LLM) can potentially be used to determine edge
orientations. We provide the implementation of this approach as both a website
and in the Python package <redacted>.
\end{abstract}
